A spiral galaxy's mass distribution uncovered through lensing and dynamics

### 2016-01-28 LensingPaper_draft_v3.05.tex
Included my typo corrections until 3.3.1.

%--------------------------------------------
\begin{abstract}
We investigate the matter distribution of a spiral galaxy with counter-rotating stellar core, SDSS J1331+3628 (J1331), independently with gravitational lensing and dynamical modelling. By fitting a gravitational potential model to a quadruplet of lensing images around J1331's bulge, we tightly constrain the mass inside the Einstein radius $R_\text{ein}=(0.91\pm0.02)''(\simeq1.83\pm0.04~\text{kpc})$ to within 4\%: $M_\text{ein} = (7.8\pm0.3) \cdot 10^{10} M_\odot$. We model observed long-slit major axis stellar kinematics in J1331's central regions by finding Multi-Gaussian Expansion (MGE) models for the stellar and dark matter distribution that solve the axisymmetric Jeans equations. The lens and dynamical model are independently derived, but in very good agreement with each other around $\sim R_\text{ein}$. We find that J1331's center requires a steep total mass-to-light ratio gradient. A dynamical model including a NFW halo (with virial velocity $v_{200} \simeq 240 \pm 40~\text{km s}^{-1}$ and concentration of $c_{200} \simeq 8 \pm 2$) and moderate tangential velocity anisotropy ($\beta_z \simeq -0.4 \pm 0.1$) can reproduce the signatures of J1331's counter-rotating core and predict the stellar and gas rotation curve at larger radii. However, our models do not agree with the observed velocity dispersion at large radii. We speculate that the reason could be a non-trivial re-distribution of matter due to a possible merger event in J1331's recent past.
\end{abstract}
%--------------------------------------------

@@

%--------------------------------------------
Gravitational lensing and dynamical modelling provide independent constraints on the mass distribution of galaxies. Combining them allows for valuable cross-checking opportunities to disentangle the stellar and dark matter (DM) content of galaxies.

@@

The mass profile at larger galactocentric radii can be probed by gas rotation curves that directly measure the galaxy's circular velocity profile (e.g., \citealt{1980ApJ...238..471R}). However, due to its dissipative nature, gas motions are very sensitive to disturbances by e.g., spiral arms and bars (e.g., \citealt{2004dad..book.....S}).

Because stars are dissipationless dynamical tracers and present almost everywhere in the galaxy, stellar dynamical modelling can complement mass constraints from lensing at small and gas motions at large radii. As stellar motions are complex---a bulk rotation around one principal axis combined with random motions in all coordinate directions---, full dynamical modelling of rotation, dispersion and velocity anisotropies is needed to deduce the matter distribution.

The Sloan WFC Edge-on Late-type Lens Survey (SWELLS, WFC = Wide field camera) \citep{SWELLSI,SWELLSII,SWELLSIII,SWELLSIV,SWELLSV,SWELLSVI} is dedicated to find and investigate spiral galaxies, which are (i) strong gravitational lenses and (ii) seen almost edge-on, such that rotation curves can be easily measured. By combining lensing and dynamical modelling degeneracies inherent in both methods can be broken.

@@

\citet{SWELLSI} confirmed that J1331 was a strong gravitational lens, measured its apparent brightness and estimated the stellar masses of disk and bulge. The lensing properties of J1331 were first analysed by \citet{SWELLSIII}. \citet{SWELLSV} measured the gas and stellar kinematics along the major axis and deduced J1331's mass profile from the gas rotation curve at large radii and total mass inside the Einstein radius from gravitational lensing, focusing mostly on the outer regions of J1331.

The goal of this work is to complement the previous work on J1331 by an in-depth analysis of the matter distribution in J1331's inner regions. We use stellar dynamical modelling in addition to lensing constraints, similar to a study of the Einstein Cross by \citet{GlennEC}. We attempt to disentangle the stellar and DM components and test if employed axisymmetric Jeans models work also in the presence of a counter-rotating core. Ideally, this work on J1331 could also help understanding how mergers modify the mass distribution of a galaxy.

This paper is organized as follows: Section \ref{sec:data} summarizes the data, Section \ref{sec:Modelling} gives an overview of the modelling techniques used in this work, and Section \ref{sec:Results} presents our results on the surface photometry of J1331 using Multi-Gaussian expansions (Section \ref{sec:MGE_results}), constraints from lensing (Section \ref{sec:results_lensing}) and Jeans modelling based on the surface brightness only (Section \ref{sec:results_JAM_SB}) and including a NFW DM halo (Section \ref{sec:results_JAM_NFW}). Finally Section \ref{sec:Discussion} uses the result to discuss J1331's stellar mass-to-light ratio, possible merger history and starting points for future work.
%--------------------------------------------

%============================================
\begin{figure*}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/first_glimpse_450.ps}
  \caption{J1331 in the F450W filter.}
  \label{fig:F450W}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/first_glimpse_814.ps}
  \caption{J1331 in the F814W filter.}
  \label{fig:F814W}
\end{subfigure}
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/lens_imgpos.ps}
  \caption{Lensing images around the galaxy center.}
  \label{fig:lens_just_imgpos}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/stellar_kinematics_data.ps}
  \caption{Stellar kinematics by \citet{SWELLSV}.}
  \label{fig:kinematics}
\end{subfigure}
\caption{Hubble Space telescope (HST) images and stellar kinematics of the galaxy SDSS J1331+3628 (J1331), which has a large counter-rotating core and whose bulge acts as a strong lens for a bluish background source. \emph{Panel (a) and (b):} HST/WFPC2/WFC3 images of J1331 by \citet{SWELLSI} in two filters, F450W in panel (a) and F814W in panel (b). The black solid line in panel (b) shows the orientation of the major-axis. Its length is $10''$ and it marks approximately the region within which we carry out the dynamical modelling. \emph{Panel (c):} Lensing images in the central region of J1331.  An IRAF \emph{ellipse} model was fitted to and then subtracted from the galaxy's F450W surface brightness in panel (a). The (smoothed) residuals within the white square in panel (a) are shown in panel (c). The four bright blobs (A, B, C and D) which are visible in the residuals are arranged in a typical strong lensing configuration around the center of the galaxy (G). (The configuration of the two additional blobs which lie approximately on a line with A, B and G does not suggest that these blobs are a lensing doublet. They might rather be star forming regions of a background galaxy.) \emph{Panel (d):} Stellar Kinematics along the galaxy's major axis as measured by \citet{SWELLSV}. Shown are line-of-sight rotation velocity $v_\text{rot}$ (top), line-of-sight velocity dispersion $\sigma$ (middle) and the root mean-square (rms) velocity $v_\text{rms} = \sqrt{v_\text{rot}^2 + \sigma^2}$ (bottom). The dashed line indicates the galaxy's effective half-light radius (in the F814W filter), $R_\text{eff} = 2.6'' = 5.2~\text{kpc}$. The $v_\text{rot}$ curve reveals that J1331 has a counter-rotating core within $R_\text{eff}$.}
\label{fig:specialJ1331}
\end{figure*}
%============================================

@@

%--------------------------------------------
\subsection{Stellar kinematics}
%--------------------------------------------

%--------------------------------------------
For the dynamical modelling we use the stellar kinematics for J1331 measured by \citet{SWELLSV}. They obtained long-slit spectra along J1331's major-axis with the Low Resolution Imaging Spectrograph (LRIS) on the Keck I 10m telescope. The width of the slit was $1''$ and the seeing conditions had a FWHM of $\sim 1.1''$. 1D spectra for spatial bins of different widths along the major axis were extracted. They measured line-of-sight stellar rotation velocities ($v_\text{rot}$) and stellar velocity dispersions ($\sigma$) from the spectral region containing the absorption lines Mgb (5177 \AA) and FeII (5270, 5335 and 5406 \AA) (analogous to \citealt{SWELLSII}). Gas kinematics were extracted from Gaussian line profile fits to H$\alpha$ (6563 \AA) and [NII] (6583 and 6548 \AA) emission lines as tracers for ionized gas.

The stellar kinematics, $v_\text{rot}$, $\sigma$ and the root-mean-square velocity $v_\text{rms}^2=v_\text{rot}^2 + \sigma^2$ are shown in Figure \ref{fig:kinematics}. The rotation curve reveals a counter-rotating core within $2''\hat{=}4~\text{kpc}$. Outside of $\sim 3.5''$ there is a drop in the dispersion. This could indicate the boundary between the pressure supported bulge and the rotationally supported disk, which appears around this radius in the F450W filter in Figure \ref{fig:F450W}. However, in the brighter F814W filter in Figure \ref{fig:F814W} the large reddish bulge extends out to $\sim5''$.

Inside of $\sim 4''$ the data appears to be symmetric; outside of it the data suggests that the assumption of axisymmetry seems not to be valid anymore. We add $-2.3~\text{km s}^{-1}$ to the $v_\text{rot}$ to ensure $v_\text{rot}(R'=0) \sim 0$ as a possible correction term for a misjudgement of the systemic velocity. We also symmetrize the data within $4''$ and asign a minimum error of $\delta v_\text{rms} > 5~\text{km s}^{-1}$ to the $v_\text{rms}$ data. In our dynamical Jeans modelling approach, which is based on the assumption of axisymmetry, only stellar kinematics with either $R'  \lesssim 2.5''$ or $R' \lesssim 4''$ are used. Another reason to restrict to modelling to the bulge region is that our surface brightness model (Table \ref{tab:MGEF814W} and Figure \ref{fig:MGEinnerRegions}) is only a good representation of J1331's F814W light distribution inside $\sim 5''$.
%--------------------------------------------

@@

%--------------------------------------------
MGEs are used to parametrize the observed surface brightness (or projected total mass) of a galaxy as a sum of $N$ 2-dimensional, elliptical Gaussians \citep{1991ApJ...366..599B,1992A&A...253..366M,1994A&A...285..723E,1999MNRAS.303..495E}. This work makes use of the algorithm and code\footnote{The IDL code package for fitting MGEs to images by \citet{Cap02} is available online at \url{http://www-astro.physics.ox.ac.uk/~mxc/software}. The version from June 2012 was used in this work.} by \citet{Cap02}. We assume all Gaussians to have the same center and position angle $\phi$, i.e., the orientation of the Gaussians' major axis measured counter-clockwise from the $y'$-axis of the coordinate system with polar coordinates $(R',\theta')$. Then the surface brightness can be written as
\begin{eqnarray}
I(R',\theta') &=& \sum_{i=1}^{N} I_{0,i} \exp\left[ - \frac{1}{2\sigma_i^2} \left(x^{'2} + \frac{y^{'2}}{q_i^{'2}}\right)\right]\label{eq:MGEgeneral}\\
\text{with } I_{0,i} &=& \frac{L_i}{2\pi \sigma_i^2 q'_i}\label{eq:centralItotalL}\\
\text{and } x'_i &=& R' \cos(\theta' - \phi)\nonumber\\
y'_i &=& R' \sin(\theta' - \phi),\nonumber
\end{eqnarray}
where $I_{0,i}$ is the central surface brightness of each Gaussian, $L_i$ its total luminosity, $\sigma_i$ its dispersion along the major axis and $q'_i$ the axis ratio between the elliptical Gaussian's major and minor axis.

\subsubsection{Point spread function (PSF)}

We can also expand the telescope's PSF as a sum of circular Gaussians,
\begin{equation}
\text{PSF}(x',y') = \sum_j \frac{G_j}{2 \pi \delta_j^2} \exp\left[- \frac{1}{2 \delta_j^2} \left({x'}^2 + {y'}^2 \right)\right], \label{eq:PSFgeneral}
\end{equation}
where $\sum_j G_j = 1$ and $\delta_j$ are in this case the dispersions of the circular PSF Gaussians.

@@

$I(R',\theta')$ describes the intrinsic, to 2D projected light distribution or surface density of the galaxy. Under the assumption that the galaxy is oblate and axisymmetric, and given the inclination angle $i$ of the galaxy with respect to the observer, MGEs allow an analytic deprojection of the 2D MGE to get a 3D light distribution or density $\nu(R,z)$ for the galaxy,
\begin{equation}
\nu(R,z) = \sum_i \nu_{0,i} \exp \left[-\frac{1}{2\sigma_i}\left(R^2 + \frac{z^2}{q_i^2} \right) \right]. \label{eq:deprojMGE}
\end{equation}
The flattening of each axisymmetric 3D Gaussian $q_i$ and its central density $\nu_{0,i}$ follow from the observed 2D axis ratio $q'_i$ and surface density $I_{0,i}$ as
\begin{eqnarray*}
q_i^2 &=& \frac{q_i'^2 - \cos^2 i}{\sin^2 i}\\
\nu_{0,i} &=& \frac{q_i' I_{0,i}}{q_i \sqrt{2 \pi \sigma_i^2}}.
\end{eqnarray*}
%--------------------------------------------

@@

%--------------------------------------------
\subsubsection{Lensing formalism}
%--------------------------------------------

%--------------------------------------------
A gravitational lens is a mass distribution, whose gravitational potential $\Phi$ acts as a lens for light coming from a source positioned somewhere on a plane behind the lens. The angular diameter distance from the observer to the lens is $D_d$, to the source plane $D_s$, and the distance between the lens and source plane is $D_{ds}$. The deflection potential of the lens is its potential, projected along the line of sight $z$ and rescaled to
\begin{equation}
\psi(\vect{\theta}) := \frac{D_{ds}}{D_d D_s} \frac{2}{c^2} \int \Phi(\vect{r}=D_d \vect{\theta},z) {\ \mathrm d} z, \label{eq:psidef}
\end{equation}
where $\vect{\theta}$ is a 2-dimensional vector on the plane of the sky. The light from the source at $\vect{\beta} = (\xi,\eta)$\footnote{$\xi$ and $\eta$ are cartesian coordinates on the plane of the sky.} is deflected according to the lens equation
\begin{equation}
\vect{\beta} = \vect{\theta}_i - \left.\vect{\nabla}_\theta \psi(\vect{\theta})\right|_{\vect{\theta}_i} \label{eq:lenseqpot}
\end{equation}
into an image $\vect{\theta}_i = (x_i,y_i)$. The gradient of the deflection potential $\vect{\nabla}_\theta \psi(\vect{\theta})$ is equal to the angle by which the light is deflected multiplied by $D_{ds}/D_{s}$.

The total time delay of an deflected light path through $\vect{\theta}$ with respect to the unperturbed light path is given by 
\begin{equation}
\Delta t(\vect{\theta}) = \frac{(1+z_d)}{c} \frac{D_d D_s}{D_{ds}} \left[ \frac 12 (\vect{\theta} - \vect{\beta})^2 - \psi(\vect{\theta})\right] \label{eq:timedelay}
\end{equation}
\citep{BartGravLens}. According to Fermat's principle the image positions will be observed at the extrema of $\Delta t(\vect{\theta})$.

The inverse magnification tensor
\begin{equation}
\mathscr{M}^{-1} \equiv \frac{\partial \vect{\beta}}{\partial \vect{\theta}} \overset{(\ref{eq:lenseqpot})}{=} \left(\delta_{ij} - \frac{\partial^2 \psi}{\partial \theta_i \partial \theta_j} \right)\label{eq:magnificationtensor}
\end{equation}
describes how the source position changes with image position. It also describes the distortion of the image shape for an extended source and its magnification due to lensing according to
$$\mu \equiv \frac{\text{image area}}{\text{source area}} = \det \mathscr{M}.$$
Lines in the image plane for which the magnification becomes infinite, i.e., $\det \mathscr{M}^{-1} = 0$, are called critical curves. The corresponding lines in the source plane are called caustics. The position of the source with respect to the caustic detemines the number of images and their configuration and shape with respect to each other.

The Einstein mass $M_\text{ein}$ and Einstein radius $R_\text{ein}$ are defined via the relation
\begin{equation*}
M_\text{ein} \equiv M_\text{proj}(<R_\text{ein}) \overset{!}{=} \pi \Sigma_\text{crit} R_\text{ein}^2,
\end{equation*}
where $\Sigma_\text{crit} \equiv \frac{c^2}{4\pi G} \frac{D_s}{D_d D_{ds}}$ is the critical density and $M_\text{proj}(<R_\text{ein})$ is the mass projected along the line-of-sight within $R_\text{ein}$. $M_\text{ein}$ is similar to the projected mass within the critical curve $M_\text{crit}$.

@@

To estimate how the uncertainties in the determination of image positions and galaxy center affect the results, we sample random positions from two-dimensional Normal distributions with peaks and standard deviations according to Table \ref{tab:lenspos}. Model fits to many sampled image positions lead to probability distributions for the best fit shape parameters and Einstein quantities; peak and standard deviations are given in the second column of Table \ref{tab:bestfitlensmodel}. We constrain the Einstein radius to within 2\%, $R_\text{ein} = (0.91 \pm 0.02)''$ and the projected mass within the critical curve with a relative error of 4\%, $M_\text{crit} =(7.9\pm0.3)\cdot 10^{10} M_\odot$. Our measurement of $R_\text{ein}$ is consistent with that from \citet{SWELLSIII}, $R_\text{ein,SWELLS} = (0.96 \pm 0.04)''$ (which used a singular isothermal ellipsoid as lens mass model and the intermediate-axis convention of the critical curve as the Einstein radius). The relative difference between our critical mass and that of \citet{SWELLSIII}, $M_\text{crit,SWELLS} =(8.86\pm0.61)\cdot 10^{10} M_\odot$, is 13\%.
%--------------------------------------------

@@

Two-dimensional stellar kinematics would greatly help to resolve this problem. The main future work would be therefore getting high-resolution integral-field spectroscopy for J1331. High spatial resolution would be required to clearly identify J1331's presumably complex kinematic structure. High spectral resolution would be important to be able to reliably measure the low velocity dispersion in the outer regions of J1331.

>>>>

%--------------------------------------------
\begin{abstract}
We investigate the matter distribution of a spiral galaxy with counter-rotating stellar core, SDSS J1331+3628 (J1331), independently with gravitational lensing and stellar dynamical modelling. By fitting a gravitational potential model to a quadruplet of lensing images around J1331's bulge, we tightly constrain the mass inside the Einstein radius $R_\text{ein}=(0.91\pm0.02)''(\simeq1.83\pm0.04~\text{kpc})$ to within 4\%: $M_\text{ein} = (7.8\pm0.3) \cdot 10^{10} M_\odot$. We model observed long-slit major axis stellar kinematics in J1331's central regions by finding Multi-Gaussian Expansion (MGE) models for the stellar and dark matter distribution that solve the axisymmetric Jeans equations. The lens and dynamical model are independently derived, but in very good agreement with each other around $\sim R_\text{ein}$. We find that J1331's center requires a steep total mass-to-light ratio gradient. A dynamical model including a NFW halo (with virial velocity $v_{200} \simeq 240 \pm 40~\text{km s}^{-1}$ and concentration $c_{200} \simeq 8 \pm 2$) and moderate tangential velocity anisotropy ($\beta_z \simeq -0.4 \pm 0.1$) can reproduce the signatures of J1331's counter-rotating core and predict the stellar and gas rotation curve at larger radii. However, our models do not agree with the observed velocity dispersion at large radii. We speculate that the reason could be a non-trivial change in structure and kinematics due to a possible merger event in J1331's recent past.
\end{abstract}
%--------------------------------------------

@@

%--------------------------------------------
Gravitational lensing and dynamical modelling provide independent constraints on the mass distribution of galaxies. Combining them allows for valuable cross-checking opportunities to disentangle the galactic stellar and dark matter (DM) content.

@@

The mass profile at larger galactocentric radii can be probed by gas rotation curves that directly measure the galaxy's circular velocity profile (e.g., \citealt{1980ApJ...238..471R}). However, due to its dissipative nature gas motions are very sensitive to disturbances by, e.g., spiral arms and bars (e.g., \citealt{2004dad..book.....S}).

Because stars are dissipationless dynamical tracers and present almost everywhere in the galaxy, stellar dynamical modelling can complement mass constraints from lensing at small and gas motions at large radii. As stellar motions are complex---a bulk rotation around one principal axis combined with random motions in all coordinate directions---\citep{2008gady.book.....B}, full dynamical modelling of rotation, dispersion and velocity anisotropies is needed to deduce the matter distribution.

The Sloan WFC Edge-on Late-type Lens Survey (SWELLS, WFC = Wide field camera) \citep{SWELLSI,SWELLSII,SWELLSIII,SWELLSIV,SWELLSV,SWELLSVI} is dedicated to find and investigate spiral galaxies, which are (i) strong gravitational lenses and (ii) observed almost edge-on, such that rotation curves can be easily measured. By combining lensing and dynamical modelling, degeneracies inherent in both methods can be broken.

@@

\citet{SWELLSI} confirmed that J1331 is a strong gravitational lens, measured its apparent brightness and estimated the stellar masses of disk and bulge. The lensing properties of J1331 were first analysed by \citet{SWELLSIII}. \citet{SWELLSV} measured the gas and stellar kinematics along the major axis and deduced J1331's mass profile from the gas rotation curve at large radii and total mass inside the Einstein radius from gravitational lensing, focusing mostly on the outer regions of J1331.

The goal of this work is to complement the previous studies of J1331 by an in-depth analysis of the matter distribution in J1331's inner regions. We use stellar dynamical modelling in addition to lensing constraints, similar to a study of the Einstein Cross by \citet{GlennEC}. We attempt to disentangle the stellar and DM components and test if employed axisymmetric Jeans models work also in the presence of a counter-rotating core. Ideally, this work on J1331 could also help understanding how mergers modify the mass distribution of a galaxy.

This paper is organized as follows: Section \ref{sec:data} summarizes the data, Section \ref{sec:Modelling} gives an overview of the modelling techniques used in this work, and Section \ref{sec:Results} presents our results on the surface photometry of J1331 using Multi-Gaussian expansions (Section \ref{sec:MGE_results}), constraints from lensing (Section \ref{sec:results_lensing}) and Jeans modelling based on the surface brightness only (Section \ref{sec:results_JAM_SB}) and including a NFW DM halo (Section \ref{sec:results_JAM_NFW}). Finally, Section \ref{sec:Discussion} uses these results to discuss J1331's stellar mass-to-light ratio, central kinematics, possible merger history and starting points for future work.
%--------------------------------------------

%============================================
\begin{figure*}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/first_glimpse_450.ps}
  \caption{J1331 in the F450W filter.}
  \label{fig:F450W}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/first_glimpse_814.ps}
  \caption{J1331 in the F814W filter.}
  \label{fig:F814W}
\end{subfigure}
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/lens_imgpos.ps}
  \caption{Lensing images around the galaxy center.}
  \label{fig:lens_just_imgpos}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{fig/stellar_kinematics_data.ps}
  \caption{Stellar kinematics by \citet{SWELLSV}.}
  \label{fig:kinematics}
\end{subfigure}
\caption{Hubble Space Telescope (HST) images and stellar kinematics of the galaxy SDSS J1331+3628 (J1331), which has a large counter-rotating core and whose bulge acts as a strong lens for a bluish background source. \emph{Panel (a) and (b):} HST/WFPC2/WFC3 images of J1331 by \citet{SWELLSI} in two filters, F450W in panel (a) and F814W in panel (b). The black solid line in panel (b) shows the orientation of the major axis. Its length is $10''$ and it marks approximately the region within which we carry out the dynamical modelling. \emph{Panel (c):} Lensing images in the central region of J1331.  An IRAF \emph{ellipse} model was fitted to and then subtracted from the galaxy's F450W surface brightness in panel (a). The (smoothed) residuals within the white square in panel (a) are shown in panel (c). The four bright blobs (A, B, C and D), which are visible in the residuals, are arranged in a typical strong lensing configuration around the center of the galaxy (G). (The configuration of the two additional blobs which lie approximately on a line with A, B and G does not suggest that these blobs are a lensing doublet. They might rather be star-forming regions of a background galaxy.) \emph{Panel (d):} Stellar kinematics along the galaxy's major axis as measured by \citet{SWELLSV}. Shown are line-of-sight rotation velocity $v_\text{rot}$ (top), line-of-sight velocity dispersion $\sigma$ (middle) and the root mean-square (rms) velocity $v_\text{rms} = \sqrt{v_\text{rot}^2 + \sigma^2}$ (bottom). The dashed line indicates the galaxy's effective half-light radius (in the F814W filter), $R_\text{eff} = 2.6'' \hat{=} 5.2~\text{kpc}$. The $v_\text{rot}$ curve reveals that J1331 has a counter-rotating core within $R_\text{eff}$.}
\label{fig:specialJ1331}
\end{figure*}
%============================================

@@

%--------------------------------------------
\subsection{Stellar kinematics} \label{sec:data_kinematics}
%--------------------------------------------

%--------------------------------------------
For the dynamical modelling we use the stellar kinematics for J1331 measured by \citet{SWELLSV}. They obtained long-slit spectra along J1331's major-axis with the Low Resolution Imaging Spectrograph (LRIS) on the Keck I 10m telescope. The width of the slit was $1''$ and the seeing conditions had a FWHM of $\sim 1.1''$. 1D spectra for spatial bins of different widths along the major axis were extracted. Line-of-sight stellar rotation velocities ($v_\text{rot}$) and stellar velocity dispersions ($\sigma$) were determined from the spectral region containing the absorption lines Mgb (5177 \AA) and FeII (5270, 5335 and 5406 \AA) (analogous to \citealt{SWELLSII}). Gas kinematics were extracted from Gaussian line profile fits to H$\alpha$ (6563 \AA) and [NII] (6583 and 6548 \AA) emission lines as tracers for ionized gas.

The stellar kinematics, $v_\text{rot}$, $\sigma$ and the root-mean-square velocity $v_\text{rms}^2=v_\text{rot}^2 + \sigma^2$ are shown in Figure \ref{fig:kinematics}. The rotation curve reveals a counter-rotating core within $2''\hat{=}4~\text{kpc}$. Outside of $\sim 3.5''$ there is a drop in the dispersion. This could indicate the boundary between the pressure supported bulge and the rotationally supported disk, which appears around this radius in the F450W filter in Figure \ref{fig:F450W}. However, in the brighter F814W filter in Figure \ref{fig:F814W} the smooth reddish bulge extends out to $\sim5''$.

Inside of $\sim 4''$ the data appears to be symmetric; outside of it the data suggests that the assumption of axisymmetry seems not to be valid anymore. We subtract $2.3~\text{km s}^{-1}$ from $v_\text{rot}$ to ensure $v_\text{rot}(R'=0) \sim 0$ as a possible correction term for a misjudgement of the systemic velocity. We also symmetrize the data within $4''$ and asign a minimum error of $\delta v_\text{rms} \geq 5~\text{km s}^{-1}$ to the $v_\text{rms}$ data. In our dynamical Jeans modelling approach, which is based on the assumption of axisymmetry, only stellar kinematics with either $R'  \lesssim 3.5''$ or $R' \lesssim 4''$ are used. Another reason to restrict the modelling to the bulge region is that our surface brightness model (Table \ref{tab:MGEF814W} and Figure \ref{fig:MGEinnerRegions}) is only a good representation of J1331's F814W light distribution inside $\sim 5''$.
%--------------------------------------------

@@

%--------------------------------------------
MGEs are used to parametrize the observed surface brightness (or projected total mass) of a galaxy as a sum of $N$ 2-dimensional, elliptical Gaussians \citep{1991ApJ...366..599B,1992A&A...253..366M,1994A&A...285..723E,1999MNRAS.303..495E}. This work makes use of the algorithm and code\footnote{The IDL code package for fitting MGEs to images by \citet{Cap02} is available online at \url{http://www-astro.physics.ox.ac.uk/~mxc/software}. The version from June 2012 was used in this work.} by \citet{Cap02}. We assume all Gaussians to have the same center and position angle $\phi$, i.e., the orientation of the Gaussians' major axis measured from North through East in the polar coordinate system $(R',\theta')$ on the plane of the sky. Then a surface brightness model can be expressed as
\begin{eqnarray}
I(R',\theta') &=& \sum_{i=1}^{N} I_{0,i} \exp\left[ - \frac{1}{2\sigma_i^2} \left(x^{'2} + \frac{y^{'2}}{q_i^{'2}}\right)\right]\label{eq:MGEgeneral}\\
\text{with } I_{0,i} &=& \frac{L_i}{2\pi \sigma_i^2 q'_i}\label{eq:centralItotalL}\\
\text{and } x'_i &=& R' \cos(\theta' - \phi)\nonumber\\
y'_i &=& R' \sin(\theta' - \phi),\nonumber
\end{eqnarray}
where $I_{0,i}$ is the central surface brightness of each Gaussian, $L_i$ its total luminosity, $\sigma_i$ its dispersion along the major axis and $q'_i$ the axis ratio between the elliptical Gaussian's major and minor axis.

\subsubsection{Convolution with the point spread function (PSF)}

We can also expand the telescope's PSF as a sum of circular Gaussians,
\begin{equation}
\text{PSF}(x',y') = \sum_j \frac{G_j}{2 \pi \delta_j^2} \exp\left[- \frac{1}{2 \delta_j^2} \left({x'}^2 + {y'}^2 \right)\right], \label{eq:PSFgeneral}
\end{equation}
where $\sum_j G_j = 1$ and $\delta_j$ are the dispersions of the circular PSF Gaussians.

@@

$I(R',\theta')$ describes the intrinsic, to 2D projected light distribution or surface density of the galaxy. Under the assumption that the galaxy is oblate and axisymmetric, and given the inclination angle $i$ of the galaxy with respect to the observer, MGEs allow an analytic deprojection of the 2-dimensional MGE to get a 3-dimensional axisymmetric light distribution or density $\nu(R,z)$ for the galaxy,
\begin{equation}
\nu(R,z) = \sum_i \nu_{0,i} \exp \left[-\frac{1}{2\sigma_i}\left(R^2 + \frac{z^2}{q_i^2} \right) \right], \label{eq:deprojMGE}
\end{equation}
where $R$ is the distance from the galaxy's short axis and $z$ the height above the galactic plane. The flattening $q_i$ of each axisymmetric 3D Gaussian and its central density $\nu_{0,i}$ follow from the observed 2D axis ratio $q'_i$ and surface density $I_{0,i}$ as
\begin{eqnarray*}
q_i^2 &=& \frac{q_i'^2 - \cos^2 i}{\sin^2 i}\\
\nu_{0,i} &=& \frac{q_i' I_{0,i}}{q_i \sqrt{2 \pi \sigma_i^2}}.
\end{eqnarray*}
%--------------------------------------------

@@

%--------------------------------------------
\subsubsection{Strong lensing formalism}
%--------------------------------------------

%--------------------------------------------
A gravitational lens is a mass distribution whose gravitational potential $\Phi$ acts as a lens for light coming from a source positioned somewhere on a plane behind the lens. We summarize the basic formalism (see also \citealt{1992grle.book.....S,1996astro.ph..6001N,2006glsw.conf....1S,2006glsw.conf...91K,2010ARA&A..48...87T}). The angular diameter distance from the observer to the lens is $D_d$, to the source plane $D_s$, and the distance between the lens and source plane is $D_{ds}$. The deflection potential of the lens is its potential projected along the line of sight $z'$ and rescaled to
\begin{equation}
\psi(\vect{\theta}) := \frac{D_{ds}}{D_d D_s} \frac{2}{c^2} \int \Phi(\vect{r}=D_d \vect{\theta},z') {\ \mathrm d} z', \label{eq:psidef}
\end{equation}
where $\vect{\theta}\equiv(x',y')$ is a 2-dimensional vector on the plane of the sky. The light from the source at $\vect{\beta} \equiv (x_s',y_s')$ is deflected according to the lens equation
\begin{equation}
\vect{\beta} = \vect{\theta}_i - \left.\vect{\nabla}_\theta \psi(\vect{\theta})\right|_{\vect{\theta}_i} \label{eq:lenseqpot}
\end{equation}
into an image $\vect{\theta}_i = (x_i',y_i')$. The gradient of the deflection potential, $\vect{\nabla}_\theta \psi(\vect{\theta})$, is equal to the angle by which the light is deflected multiplied by $D_{ds}/D_{s}$.

The total time delay of a deflected light path through $\vect{\theta}$ with respect to the unperturbed light path is given by 
\begin{equation}
\Delta t(\vect{\theta}) = \frac{(1+z_d)}{c} \frac{D_d D_s}{D_{ds}} \left[ \frac 12 (\vect{\theta} - \vect{\beta})^2 - \psi(\vect{\theta})\right]. \label{eq:timedelay}
\end{equation}
 According to Fermat's principle the image positions will be observed at the extrema of $\Delta t(\vect{\theta})$.

The inverse magnification tensor
\begin{equation}
\mathscr{M}^{-1} \equiv \frac{\partial \vect{\beta}}{\partial \vect{\theta}} \overset{(\ref{eq:lenseqpot})}{=} \left(\delta_{ij} - \frac{\partial^2 \psi}{\partial \theta_i \partial \theta_j} \right)\label{eq:magnificationtensor}
\end{equation}
describes how the source position changes with image position. It also describes the distortion of the image shape for an extended source and its magnification $\mu$ due to lensing according to
$$\mu \equiv \frac{\text{image area}}{\text{source area}} = \det \mathscr{M}.$$
Lines in the image plane for which the magnification becomes infinite, i.e., $\det \mathscr{M}^{-1} = 0$, are called critical curves. The corresponding lines in the source plane are called caustics. The position of the source with respect to the caustic detemines the number of images and their configuration and shape with respect to each other.

The Einstein mass $M_\text{ein}$ and Einstein radius $R_\text{ein}$ are defined via the relation
\begin{equation*}
M_\text{ein} \equiv M(<R_\text{ein}) \overset{!}{=} \pi \Sigma_\text{crit} R_\text{ein}^2,
\end{equation*}
where $\Sigma_\text{crit} \equiv \frac{c^2}{4\pi G} \frac{D_s}{D_d D_{ds}}$ is the critical density and $M(<R_\text{ein})$ is the mass projected along the line-of-sight within $R'=R_\text{ein}$. $M_\text{ein}$ is similar to the projected mass within the critical curve $M_\text{crit}$.

@@

To estimate how the uncertainties in the determination of image positions and galaxy center affect the results, we sample random positions from 2-dimensional Normal distributions with peaks and standard deviations according to Table \ref{tab:lenspos}. Model fits to many sampled image positions lead to probability distributions for the best fit shape parameters and Einstein quantities; peak and standard deviations are given in the second column of Table \ref{tab:bestfitlensmodel}. We constrain the Einstein radius to within 2\%, $R_\text{ein} = (0.91 \pm 0.02)''$ and the projected mass within the critical curve with a relative error of 4\%, $M_\text{crit} =(7.9\pm0.3)\cdot 10^{10} M_\odot$. Our measurement of $R_\text{ein}$ is consistent with that from \citet{SWELLSIII}, $R_\text{ein,SWELLS} = (0.96 \pm 0.04)''$ (which used a singular isothermal ellipsoid as lens mass model and the intermediate-axis convention of the critical curve as the Einstein radius). The relative difference between our critical mass and that of \citet{SWELLSIII}, $M_\text{crit,SWELLS} =(8.86\pm0.61)\cdot 10^{10} M_\odot$, is 13\%.
%--------------------------------------------

@@

2-dimensional stellar kinematics would greatly help to resolve this problem. The main future work would be therefore getting high-resolution integral-field spectroscopy for J1331. High spatial resolution would be required to clearly identify J1331's presumably complex kinematic structure. High spectral resolution would be important to be able to reliably measure the low velocity dispersion in the outer regions of J1331.
